---
title: "R Notebook"
output: html_notebook
---

```{r}
library(readr)
library(tidyverse)
```

```{r}
GameResults <- read_csv("newestData.csv")
```


# Developing a model for Spread


```{r}
head(GameResults)
```

```{r}
GameResults %>% group_by(Year) %>%
  summarize(mean(Spread))
```

```{r}
SpreadMatrix <- GameResults %>%
  dplyr::select(-c(2:4,10:11,113:114,116:125,127:128,130:139)) %>%
  dplyr::select(Spread, everything())
```

```{r}
hist(SpreadMatrix$Spread)

# spread fairly normally distributed
```

```{r}
library(leaps)
set.seed(123)
train <- sample.int(5324, size=2000)
SpreadMatrix.train <- SpreadMatrix[train,]
SpreadMatrix.test <- SpreadMatrix[-train,]

#regfit.best=regsubsets(Spread~.,data=SpreadMatrix.train, nvmax =19)
#test.mat=model.matrix(Spread∼.,data=SpreadMatrix.test)


```

```{r}
# Backward Elimination 
#Full=lm(Spread~., data=SpreadMatrix.train)
#summary(Full)
#MSE=(summary(Full)$sigma)^2
#step(Full, scale=MSE, trace=0)
```


lm(formula = Spread ~ stadium_neutral + Vegas_Pred_Spread + Off_Total_Points_home + 
    Off_Total_Yards_home + Off_Yards_Per_Attempt_home + Off_Total1stD_home + 
    Off_Passing_Yds_home + `Off_Passing Y/A_home` + Off_Rushing_TD_home + 
    Off_Penalties_home + Off_Expected_Points_home + Def_Total_Points_home + 
    Def_Turnovers_home + `Def_TO%_home` + Def_Expected_Points_home + 
    Off_Yards_Per_Attempt_away + Off_Turnovers_away + Off_Passing_Yds_away + 
    Off_Passing_TD_away + `Off_Passing Y/A_away` + Off_Passing_1stD_away + 
    `Off_Rushing Y/A_away` + `Off_Sc%_away` + `Off_TO%_away` + 
    Def_Total_Points_away + Def_Turnovers_away + Def_Passing_TD_away + 
    Def_Rushing_TD_away + `Def_TO%_away` + RZPct_home + `3D%_away` + 
    RZPct_away + PB_Rank_home, data = SpreadMatrix.train)
    

```{r}
# single backward elimination on training data
BE.1 <- lm(formula = Spread ~ stadium_neutral + Vegas_Pred_Spread + Off_Total_Points_home + 
    Off_Total_Yards_home + Off_Yards_Per_Attempt_home + Off_Total1stD_home + 
    Off_Passing_Yds_home + `Off_Passing Y/A_home` + Off_Rushing_TD_home + 
    Off_Penalties_home + Off_Expected_Points_home + Def_Total_Points_home + 
    Def_Turnovers_home + `Def_TO%_home` + Def_Expected_Points_home + 
    Off_Yards_Per_Attempt_away + Off_Turnovers_away + Off_Passing_Yds_away + 
    Off_Passing_TD_away + `Off_Passing Y/A_away` + Off_Passing_1stD_away + 
    `Off_Rushing Y/A_away` + `Off_Sc%_away` + `Off_TO%_away` + 
    Def_Total_Points_away + Def_Turnovers_away + Def_Passing_TD_away + 
    Def_Rushing_TD_away + `Def_TO%_away` + RZPct_home + `3D%_away` + 
    RZPct_away + PB_Rank_home, data = SpreadMatrix.train)


pred <- (SpreadMatrix.test$Spread - predict(BE.1, newdata=SpreadMatrix.test))^2

sqrt(mean(pred))
```


# Backward Selection Attempt I

```{r}
library(MASS)
library(caret)

# just trying to fit on all 

# Set seed for reproducibility
set.seed(123)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.model <- train(Spread ~., data = SpreadMatrix,
                    method = "leapBackward", 
                    tuneGrid = data.frame(nvmax = 1:80),
                    trControl = train.control
                    )
step.model$results
```

```{r}
which.min(step.model$results$RMSE)
min(step.model$results$RMSE)
```

```{r}
plot(step.model$results$RMSE)
```


```{r}
library(plm)



SpreadMatrix.1 <- SpreadMatrix %>%
  dplyr::select(-c(5:6))

detect.lindep(SpreadMatrix.1)
```


```{r}
sort(abs(cor(SpreadMatrix.1)[,1]), decreasing=TRUE)

# Vegas_Pred_Spread, Off_Rank_home, Off_Total_Points_home, Off_Sc%_home, Off_Passing Y/A_home, Def_Rank_home, Def_Total_Points_home, Def_Sc%_home, Def_Passing_Y/A_home, TurnoverRatio_home, 3D%_home, Off_Passing_TD_home, Off_Rushing_TD_home, Off_Turnovers_home, Def_Passing_TD_home, Def_Rushing_TD_home, Def_Turnovers_home, Off_Total_1st_home,Off_Expected_Points_home, Off_Total_Yards_home, Def_Total_Yards_home, RB_Rank_home 
```

# Backward Selection Attempt II

```{r}
SpreadMatrix.2 <- SpreadMatrix.1 %>%
  dplyr::select(-c(8, 10, 11, 12, 14, 16, 18, 19, 22, 25, 31, 33, 34, 35, 37, 39, 41, 42, 45, 48, 54, 56, 57, 58, 60, 62, 64, 65, 68, 71, 77, 79, 80, 81, 83, 85, 87, 88, 91, 94))

# removing all linear dependencies and fitting 

set.seed(123)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.model.2 <- train(Spread ~., data = SpreadMatrix.2,
                    method = "leapBackward", 
                    tuneGrid = data.frame(nvmax = 1:50),
                    trControl = train.control
                    )
step.model.2$results

```

```{r}
which.min(step.model.2$results$RMSE)
min(step.model.2$results$RMSE)
```

```{r}
plot(step.model.2$results$RMSE)
```

```{r}
coef(step.model.2$finalModel, id=6)
```



# Backward Selection Attempt III
including some of the variables that were deemed suspicious by linear dependence function 

```{r}
SpreadMatrix.3 <- SpreadMatrix.1 %>%
  dplyr::select(-Off_Rushing_Yds_home, -Off_Passing_Yds_home,-Def_Total_Yards_home,-Off_Int_home, -Off_Fumbles_home, -Def_Int_home, -Def_Fumbles_home, -Off_Total1stD_home, - Def_Total1stD_home, -Off_Rushing_Yds_away, -Off_Passing_Yds_away,-Def_Total_Yards_away,-Off_Int_away, -Off_Fumbles_away, -Def_Int_away, -Def_Fumbles_away, -Off_Total1stD_away, - Def_Total1stD_away)


set.seed(123)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.model.3 <- train(Spread ~., data = SpreadMatrix.3,
                    method = "leapBackward", 
                    tuneGrid = data.frame(nvmax = 1:90),
                    trControl = train.control
                    )
step.model.3$results
```

```{r}
which.min(step.model.3$results$RMSE)
min(step.model.3$results$RMSE)

#12.09638 + 2*0.2709257
```

```{r}
plot(step.model.3$results$RMSE)
```

```{r}
coef(step.model.3$finalModel, id=6)
```

# Creating lm model based off best Backward Selection Model

```{r}
games2020 <- read_csv("Total2020.csv")
#head(games2020)

games2020.1 <- games2020[1:13,] %>%
  dplyr::select(team_home, team_away, Spread,Off_Total_Points_home, Def_Total_Points_home, Def_Expected_Points_home, Off_Total_Points_away, Def_Total_Points_away)


games2020.1 <- games2020.1[-9,]


games2020.1
```

```{r}
games2020.1$Vegas_Pred_Spread = c(3.5,-6,2.5,-3,6.5,19.5,-4,3.5,-4,3,7.5,-10.5)
```

```{r}
#predict(step.model.3$method, newdata = games2020.1)

lmtest <- lm(Spread~Off_Total_Points_home+Def_Total_Points_home+Def_Expected_Points_home+Off_Total_Points_away+ Def_Total_Points_away+Vegas_Pred_Spread, data=SpreadMatrix)

games2020.1$Spread <- predict(lmtest, newdata=games2020.1)

summary(lmtest)

games2020.2 <- games2020.1[,1:3]
```

```{r}
SpreadMatrix.5 <- SpreadMatrix %>%
  dplyr::select(Spread,Off_Total_Points_home,Def_Total_Points_home,Def_Expected_Points_home,Off_Total_Points_away, Def_Total_Points_away,Vegas_Pred_Spread)

set.seed(123)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.model.3.1 <- train(Spread ~.*., data = SpreadMatrix.5,
                    method = "leapBackward", 
                    tuneGrid = data.frame(nvmax = 1:20),
                    trControl = train.control
                    )
step.model.3.1$results
```

```{r}
which.min(step.model.3.1$results$RMSE)
min(step.model.3.1$results$RMSE)

#12.10237 + 2*0.2434441
```

```{r}
coef(step.model.3.1$finalModel, id=10)



```

```{r}
lmtest2 <- lm(Spread~Def_Total_Points_home + Def_Expected_Points_home + Vegas_Pred_Spread + Off_Total_Points_home*Def_Total_Points_away + Def_Total_Points_home*Off_Total_Points_away + Def_Total_Points_home*Def_Total_Points_away + Def_Total_Points_home*Vegas_Pred_Spread +
Def_Expected_Points_home*Def_Total_Points_away + Off_Total_Points_away*Def_Total_Points_away + Def_Total_Points_away*Vegas_Pred_Spread, data=SpreadMatrix) 

games2020.1 <- games2020[1:13,] %>%
  dplyr::select(team_home, team_away, Spread,Off_Total_Points_home, Def_Total_Points_home, Def_Expected_Points_home, Off_Total_Points_away, Def_Total_Points_away)


games2020.1 <- games2020.1[-9,]


games2020.1

games2020.1$Vegas_Pred_Spread = c(3.5,-6,2.5,-3,6.5,19.5,-4,3.5,-4,3,7.5,-10.5)

#summary(lmtest2)

# look into adding interaction between each offense and defense and revalidating 
                      
```


# Retrying Best Model With Addition of Important Interaction Effects

```{r}
SpreadMatrix.removed <- SpreadMatrix[SpreadMatrix$Spread < 45 & SpreadMatrix$Spread > -45,]

set.seed(123) 
train.control <- trainControl(method = "cv", number = 10)
# Train the model
model <- train(Spread~Off_Total_Points_home+Def_Total_Points_home+Def_Expected_Points_home+Off_Total_Points_away+ Def_Total_Points_away+Vegas_Pred_Spread+Off_Total_Points_home*Def_Total_Points_away+Off_Total_Points_away*Def_Total_Points_home, data = SpreadMatrix.removed, method = "lm",
               trControl = train.control)
# Summarize the results
print(model)

model$results
```

```{r}
summary(model)
```

```{r}
gamesFinalPredicts <- games2020.1[,1:3]
gamesFinalPredicts$Spread <- predict(model, newdata=games2020.1)
```


# Trying with different variables and interaction terms

```{r}
set.seed(123) 
train.control <- trainControl(method = "cv", number = 10)
# Train the model
model.1 <- train(Spread~Off_Total_Points_home+Def_Total_Points_home+Def_Expected_Points_home+Off_Total_Points_away+ Def_Total_Points_away+Vegas_Pred_Spread+Off_Total_Points_home*Def_Total_Points_away+Off_Total_Points_away*Def_Total_Points_home, data = SpreadMatrix, method = "lm",
               trControl = train.control)
# Summarize the results
print(model.1)

model.1$results
```

```{r}
SpreadMatrix$Spread.hat1 = predict(model, newdata=SpreadMatrix)
ggplot(SpreadMatrix) +
  geom_point(aes(x=Spread,y=Spread.hat1),color="lightskyblue2") +
  geom_abline(a=0,b=1,linetype="dashed") +
  theme_minimal() +
  ylab("Predicted Spread") +
  xlab("Actual Spread") +
  ggtitle("Predicted Spread v. Actual Spread for Backwards Elimination Model")


ggplot(SpreadMatrix) +
  geom_histogram(aes(x=Spread-Spread.hat1),fill="lightskyblue2") +
  theme_minimal() +
  xlab("Residuals") +
  ylab("Frequency")
```


# Forward Selection Attempt I

```{r}
# Set seed for reproducibility
set.seed(123)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.model.4 <- train(Spread ~., data = SpreadMatrix,
                    method = "leapForward", 
                    tuneGrid = data.frame(nvmax = 1:80),
                    trControl = train.control
                    )
step.model.4$results
```

```{r}
which.min(step.model.4$results$RMSE)
min(step.model.4$results$RMSE)
```

# Forward Selection Attempt II

```{r}
set.seed(123)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.model.5 <- train(Spread ~., data = SpreadMatrix.2,
                    method = "leapForward", 
                    tuneGrid = data.frame(nvmax = 1:50),
                    trControl = train.control
                    )
step.model.5$results
```

```{r}
which.min(step.model.5$results$RMSE)
min(step.model.5$results$RMSE)
```








# Forward Selection Attempt III

```{r}
set.seed(123)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.model.6 <- train(Spread ~., data = SpreadMatrix.3,
                    method = "leapForward", 
                    tuneGrid = data.frame(nvmax = 1:90),
                    trControl = train.control
                    )
step.model.6$results
```

```{r}
which.min(step.model.6$results$RMSE)
min(step.model.6$results$RMSE)
```

- backward selection outperformed forward selection 


# Stepwise Regression I

```{r}
set.seed(123)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.model.7 <- train(Spread ~., data = SpreadMatrix.2,
                    method = "leapSeq", 
                    tuneGrid = data.frame(nvmax = 1:50),
                    trControl = train.control
                    )
step.model.7$results
```

```{r}
which.min(step.model.7$results$RMSE)
min(step.model.7$results$RMSE)
```

slightly bigger than Backward Regression III, but smaller standard deviation

```{r}
coef(step.model.7$finalModel, id=6)
```


# Stepwise Regression II

```{r}
set.seed(123)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.model.8 <- train(Spread ~., data = SpreadMatrix.3,
                    method = "leapSeq", 
                    tuneGrid = data.frame(nvmax = 1:90),
                    trControl = train.control
                    )
step.model.8$results
```

```{r}
which.min(step.model.8$results$RMSE)
min(step.model.8$results$RMSE)
```

# Stepwise Regression w/o Vegas_Pred_Spread

```{r}
SpreadMatrix.6 <- SpreadMatrix.3 %>%
  dplyr::select(-Vegas_Pred_Spread)


set.seed(123)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.model.9 <- train(Spread ~., data = SpreadMatrix.6,
                    method = "leapSeq", 
                    tuneGrid = data.frame(nvmax = 1:89),
                    trControl = train.control
                    )
step.model.9$results
```

```{r}
which.min(step.model.9$results$RMSE)
min(step.model.9$results$RMSE)
```

```{r}
coef(step.model.9$finalModel, id=4)
```

```{r}
lmtest3 <- lm(Spread~Off_Total_Points_home+Def_Total_Points_home+Def_Total_Points_away+Off_Total_Points_away, data=SpreadMatrix)
summary(lmtest3)
```


# Improving on This StepWise Model

```{r}
set.seed(123) 
train.control <- trainControl(method = "cv", number = 10)
# Train the model
model.2 <- train(Spread~Off_Total_Points_home+Def_Total_Points_home+ Off_Total_Points_away+ Def_Total_Points_away+Off_Total_Points_home*Def_Total_Points_away+Off_Total_Points_away*Def_Total_Points_home, data = SpreadMatrix, method = "lm",
               trControl = train.control)
# Summarize the results
print(model.2)

model.2$results
```

```{r}
summary(model.2)
```



# Regularized Models I

```{r}
# SpreadMatrix.1 - all possible predictors & other value 
library(modelr)

SpreadMatrix.1 <- SpreadMatrix.1 %>%
  dplyr::select(-Spread.hat)

y = SpreadMatrix.1$Spread
X=model_matrix(SpreadMatrix.1,Spread~.)[,-1]
var.names=names(X)
dim(X)
```

```{r}
library(glmnet)
set.seed(216)
cvmod.0=cv.glmnet(y=y,x=as.matrix(X),type.measure = "mse",alpha=0)
set.seed(216)
cvmod.25=cv.glmnet(y=y,x=as.matrix(X), type.measure = "mse",alpha=0.25)
set.seed(216)
cvmod.5=cv.glmnet(y=y,x=as.matrix(X), type.measure = "mse",alpha=0.5)
set.seed(216)
cvmod.75=cv.glmnet(y=y,x=as.matrix(X), type.measure = "mse",alpha=0.75)
set.seed(216)
cvmod.1=cv.glmnet(y=y,x=as.matrix(X), type.measure = "mse",alpha=1)

CV.0.ERROR=cvmod.0$cvm[which(cvmod.0$lambda==cvmod.0$lambda.1se)]
CV.25.ERROR=cvmod.25$cvm[which(cvmod.25$lambda==cvmod.25$lambda.1se)]
CV.5.ERROR=cvmod.5$cvm[which(cvmod.5$lambda==cvmod.5$lambda.1se)]
CV.75.ERROR=cvmod.75$cvm[which(cvmod.75$lambda==cvmod.75$lambda.1se)]
CV.1.ERROR=cvmod.1$cvm[which(cvmod.1$lambda==cvmod.1$lambda.1se)]

MOD.RESULT=tibble(alpha=c(0,0.25,0.5,0.75,1),
                  lambda=c(cvmod.0$lambda.1se,cvmod.25$lambda.1se,
                           cvmod.5$lambda.1se,cvmod.75$lambda.1se,
                           cvmod.1$lambda.1se),
                  CV.Error=c(CV.0.ERROR,CV.25.ERROR,CV.5.ERROR,
                             CV.75.ERROR,CV.1.ERROR))
print(MOD.RESULT)
```

```{r}
best.alpha=MOD.RESULT$alpha[which.min(MOD.RESULT$CV.Error)]
best.lambda=MOD.RESULT$lambda[which.min(MOD.RESULT$CV.Error)]

best.mod=glmnet(y=y,x=as.matrix(X),nlambda=1,lambda=best.lambda,alpha=best.alpha)
best.coef=as.tibble(as.matrix(coef(best.mod)))
best.coef2=best.coef %>% 
              mutate(Parameter=c("Int",var.names)) %>%
              rename(Estimate=s0) %>%
              dplyr::select(Parameter,Estimate)
nonzero.best.coef=best.coef2 %>%
                    filter(Estimate!=0)
print(nonzero.best.coef,n=1e3)

SpreadMatrix.1$Spread.hat=predict(best.mod,newx=as.matrix(X))

ggplot(SpreadMatrix.1) +
  geom_point(aes(x=Spread,y=Spread.hat),color="lightskyblue2") +
  geom_abline(a=0,b=1,linetype="dashed") +
  theme_minimal() +
  ylab("Predicted Spread") +
  xlab("Actual Spread")

ggplot(SpreadMatrix.1) +
  geom_histogram(aes(x=Spread-Spread.hat),fill="lightskyblue2") +
  theme_minimal() +
  xlab("Residuals") +
  ylab("Frequency")
```

```{r}
sqrt(min(MOD.RESULT$CV.Error))
```


Summary of Best Models So Far:
1. Backward Selection III
2. Backward Selection II
3. Stepwise Selection I
4. GLM Net with whole data

# Regularized Models II

```{r}
# using SpreadMatrix.3
SpreadMatrix.2 <- SpreadMatrix.2 %>%
  dplyr::select(-Spread.hat)

y = SpreadMatrix.2$Spread
X=model_matrix(SpreadMatrix.2,Spread~.)[,-1]
var.names=names(X)
dim(X)
```

```{r}
set.seed(216)
cvmod.0=cv.glmnet(y=y,x=as.matrix(X),type.measure = "mse",alpha=0)
set.seed(216)
cvmod.25=cv.glmnet(y=y,x=as.matrix(X), type.measure = "mse",alpha=0.25)
set.seed(216)
cvmod.5=cv.glmnet(y=y,x=as.matrix(X), type.measure = "mse",alpha=0.5)
set.seed(216)
cvmod.75=cv.glmnet(y=y,x=as.matrix(X), type.measure = "mse",alpha=0.75)
set.seed(216)
cvmod.1=cv.glmnet(y=y,x=as.matrix(X), type.measure = "mse",alpha=1)

CV.0.ERROR=cvmod.0$cvm[which(cvmod.0$lambda==cvmod.0$lambda.1se)]
CV.25.ERROR=cvmod.25$cvm[which(cvmod.25$lambda==cvmod.25$lambda.1se)]
CV.5.ERROR=cvmod.5$cvm[which(cvmod.5$lambda==cvmod.5$lambda.1se)]
CV.75.ERROR=cvmod.75$cvm[which(cvmod.75$lambda==cvmod.75$lambda.1se)]
CV.1.ERROR=cvmod.1$cvm[which(cvmod.1$lambda==cvmod.1$lambda.1se)]

MOD.RESULT=tibble(alpha=c(0,0.25,0.5,0.75,1),
                  lambda=c(cvmod.0$lambda.1se,cvmod.25$lambda.1se,
                           cvmod.5$lambda.1se,cvmod.75$lambda.1se,
                           cvmod.1$lambda.1se),
                  CV.Error=c(CV.0.ERROR,CV.25.ERROR,CV.5.ERROR,
                             CV.75.ERROR,CV.1.ERROR))
print(MOD.RESULT)
```


```{r}
best.alpha=MOD.RESULT$alpha[which.min(MOD.RESULT$CV.Error)]
best.lambda=MOD.RESULT$lambda[which.min(MOD.RESULT$CV.Error)]

best.mod=glmnet(y=y,x=as.matrix(X),nlambda=1,lambda=best.lambda,alpha=best.alpha)
best.coef=as.tibble(as.matrix(coef(best.mod)))
best.coef2=best.coef %>% 
              mutate(Parameter=c("Int",var.names)) %>%
              rename(Estimate=s0) %>%
              dplyr::select(Parameter,Estimate)
nonzero.best.coef=best.coef2 %>%
                    filter(Estimate!=0)
print(nonzero.best.coef,n=1e3)

SpreadMatrix.2$Spread.hat=predict(best.mod,newx=as.matrix(X))

ggplot(SpreadMatrix.2) +
  geom_point(aes(x=Spread,y=Spread.hat),color="lightskyblue2") +
  geom_abline(a=0,b=1,linetype="dashed") +
  theme_minimal() +
  ylab("Predicted Spread") +
  xlab("Actual Spread")

ggplot(SpreadMatrix.2) +
  geom_histogram(aes(x=Spread-Spread.hat),fill="lightskyblue2") +
  theme_minimal() +
  xlab("Residuals") +
  ylab("Frequency")
```

```{r}
sqrt(min(MOD.RESULT$CV.Error))
```

# Regularized Models III

```{r}
SpreadMatrix.4 <- SpreadMatrix %>%
  dplyr::select(Spread,Vegas_Pred_Spread, Off_Rank_home, Off_Total_Points_home, `Off_Sc%_home`, `Off_Passing Y/A_home`, Def_Rank_home, Def_Total_Points_home, `Def_Sc%_home`, `Def_Passing Y/A_home`, TurnoverRatio_home, `3D%_home`, Off_Passing_TD_home, Off_Rushing_TD_home, Off_Turnovers_home, Def_Passing_TD_home, Def_Rushing_TD_home, Def_Turnovers_home, Off_Total1stD_home,Off_Expected_Points_home, Off_Total_Yards_home, Def_Total_Yards_home, RB_Rank_home,Off_Rank_away, Off_Total_Points_away, `Off_Sc%_away`, `Off_Passing Y/A_away`, Def_Rank_away, Def_Total_Points_away, `Def_Sc%_away`, `Def_Passing Y/A_away`, TurnoverRatio_away, `3D%_away`, Off_Passing_TD_away, Off_Rushing_TD_away, Off_Turnovers_away, Def_Passing_TD_away, Def_Rushing_TD_away, Def_Turnovers_away, Off_Total1stD_away,Off_Expected_Points_away, Off_Total_Yards_away, Def_Total_Yards_away, RB_Rank_away) %>%
  mutate(Off_Rank_diff = Off_Rank_home - Off_Rank_away,
         Off_Total_Points_diff = Off_Total_Points_home - Off_Total_Points_away,
         `Off_Sc%_diff` = `Off_Sc%_home` - `Off_Sc%_away`,
         `Off_Passing Y/A_diff` = `Off_Passing Y/A_home` - `Off_Passing Y/A_away`,
         Def_Rank_diff = Def_Rank_home - Def_Rank_away,
         Def_Total_Points_diff = Def_Total_Points_home - Def_Total_Points_away,
         `Def_Sc%_diff` = `Def_Sc%_home` - `Def_Sc%_away`,
         `Def_Passing Y/A_diff` = `Def_Passing Y/A_home` - `Def_Passing Y/A_away`,
         TurnoverRatio_diff = TurnoverRatio_home - TurnoverRatio_away,
         `3D%_diff` = `3D%_home` - `3D%_away`,
         Off_Passing_TD_diff = Off_Passing_TD_home - Off_Passing_TD_away,
         Off_Rushing_TD_diff = Off_Rushing_TD_home - Off_Rushing_TD_away,
         Off_Turnovers_diff = Off_Turnovers_home - Off_Turnovers_away,
         Def_Passing_TD_diff = Def_Passing_TD_home - Def_Passing_TD_away,
         Def_Rushing_TD_diff = Def_Rushing_TD_home - Def_Rushing_TD_away,
         Def_Turnovers_diff = Def_Turnovers_home - Def_Turnovers_away,
        Off_Total1stD_diff = Off_Total1stD_home - Off_Total1stD_away,
        Off_Expected_Points_diff = Off_Expected_Points_home - Off_Expected_Points_away, 
        Off_Total_Yards_diff = Off_Total_Yards_home - Off_Total_Yards_away, 
        Def_Total_Yards_diff = Def_Total_Yards_home - Def_Total_Yards_away, 
        RB_Rank_diff = RB_Rank_home - RB_Rank_away) %>%
  dplyr::select(-c(Off_Rank_home, Off_Total_Points_home, `Off_Sc%_home`, `Off_Passing Y/A_home`, Def_Rank_home, Def_Total_Points_home, `Def_Sc%_home`, `Def_Passing Y/A_home`, TurnoverRatio_home, `3D%_home`, Off_Passing_TD_home, Off_Rushing_TD_home, Off_Turnovers_home, Def_Passing_TD_home, Def_Rushing_TD_home, Def_Turnovers_home, Off_Total1stD_home,Off_Expected_Points_home, Off_Total_Yards_home, Def_Total_Yards_home, RB_Rank_home,Off_Rank_away, Off_Total_Points_away, `Off_Sc%_away`, `Off_Passing Y/A_away`, Def_Rank_away, Def_Total_Points_away, `Def_Sc%_away`, `Def_Passing Y/A_away`, TurnoverRatio_away, `3D%_away`, Off_Passing_TD_away, Off_Rushing_TD_away, Off_Turnovers_away, Def_Passing_TD_away, Def_Rushing_TD_away, Def_Turnovers_away, Off_Total1stD_away,Off_Expected_Points_away, Off_Total_Yards_away, Def_Total_Yards_away, RB_Rank_away))


y = SpreadMatrix.4$Spread
X=model_matrix(SpreadMatrix.4,Spread~.*.)[,-1]
var.names=names(X)
dim(X)
```

```{r}

set.seed(216)
cvmod.0=cv.glmnet(y=y,x=as.matrix(X),type.measure = "mse",alpha=0)
set.seed(216)
cvmod.25=cv.glmnet(y=y,x=as.matrix(X), type.measure = "mse",alpha=0.25)
set.seed(216)
cvmod.5=cv.glmnet(y=y,x=as.matrix(X), type.measure = "mse",alpha=0.5)
set.seed(216)
cvmod.75=cv.glmnet(y=y,x=as.matrix(X), type.measure = "mse",alpha=0.75)
set.seed(216)
cvmod.90=cv.glmnet(y=y,x=as.matrix(X), type.measure = "mse",alpha=0.9)
set.seed(216)
cvmod.1=cv.glmnet(y=y,x=as.matrix(X), type.measure = "mse",alpha=1)


CV.0.ERROR=cvmod.0$cvm[which(cvmod.0$lambda==cvmod.0$lambda.1se)]
CV.25.ERROR=cvmod.25$cvm[which(cvmod.25$lambda==cvmod.25$lambda.1se)]
CV.5.ERROR=cvmod.5$cvm[which(cvmod.5$lambda==cvmod.5$lambda.1se)]
CV.75.ERROR=cvmod.75$cvm[which(cvmod.75$lambda==cvmod.75$lambda.1se)]
CV.90.ERROR=cvmod.90$cvm[which(cvmod.90$lambda==cvmod.90$lambda.1se)]
CV.1.ERROR=cvmod.1$cvm[which(cvmod.1$lambda==cvmod.1$lambda.1se)]

MOD.RESULT=tibble(alpha=c(0,0.25,0.5,0.75,0.9,1),
                  lambda=c(cvmod.0$lambda.1se,cvmod.25$lambda.1se,
                           cvmod.5$lambda.1se,cvmod.75$lambda.1se,cvmod.90$lambda.1se,
                           cvmod.1$lambda.1se),
                  CV.Error=c(CV.0.ERROR,CV.25.ERROR,CV.5.ERROR,
                             CV.75.ERROR,CV.90.ERROR,CV.1.ERROR))
print(MOD.RESULT)
```

```{r}
best.alpha=MOD.RESULT$alpha[which.min(MOD.RESULT$CV.Error)]
best.lambda=MOD.RESULT$lambda[which.min(MOD.RESULT$CV.Error)]

best.mod=glmnet(y=y,x=as.matrix(X),nlambda=1,lambda=best.lambda,alpha=best.alpha)
best.coef=as.tibble(as.matrix(coef(best.mod)))
best.coef2=best.coef %>% 
              mutate(Parameter=c("Int",var.names)) %>%
              rename(Estimate=s0) %>%
              dplyr::select(Parameter,Estimate)
nonzero.best.coef=best.coef2 %>%
                    filter(Estimate!=0)
print(nonzero.best.coef,n=1e3)

SpreadMatrix.4$Spread.hat=predict(best.mod,newx=as.matrix(X))

ggplot(SpreadMatrix.4) +
  geom_point(aes(x=Spread,y=Spread.hat),color="lightskyblue2") +
  geom_abline(a=0,b=1,linetype="dashed") +
  theme_minimal() +
  ylab("Predicted Spread") +
  xlab("Actual Spread") +
  ggtitle("Predicted Spread v. Actual Spread for Regularized Model")

ggplot(SpreadMatrix.4) +
  geom_histogram(aes(x=Spread-Spread.hat),fill="lightskyblue2") +
  theme_minimal() +
  xlab("Residuals") +
  ylab("Frequency") + 
  ggtitle("Predicted Spread v. Actual Spread for Regularized Model")
```

```{r}
sqrt(min(MOD.RESULT$CV.Error))
```

```{r}
nonzero.best.coef
```

# Final Predictions for Weeks 2-3

```{r}
games2020predictor <- games2020[14:40,] %>%
  dplyr::select(Spread,team_home, team_away, Off_Rank_home,Off_Total_Points_home,`Off_Sc%_home`,Def_Total_Points_home, Off_Rank_away,Off_Total_Points_away,`Off_Sc%_away`,Def_Total_Points_away) %>%
  mutate(Off_Rank_diff = Off_Rank_home - Off_Rank_away,
         Off_Total_Points_diff = Off_Total_Points_home - Off_Total_Points_away,
         `Off_Sc%_diff` = `Off_Sc%_home` - `Off_Sc%_away`,
         Def_Total_Points_diff = Def_Total_Points_home - Def_Total_Points_away) %>%
  dplyr::select(Spread, team_home, team_away,Off_Rank_diff,Off_Total_Points_diff,`Off_Sc%_diff`,Def_Total_Points_diff)



for (i in 1:27) {
  games2020predictor$Spread[i] = nonzero.best.coef[1,2] + nonzero.best.coef[2,2] * games2020predictor$Off_Rank_diff[i] + nonzero.best.coef[3,2] * games2020predictor$Off_Total_Points_diff[i] + nonzero.best.coef[4,2] * games2020predictor$`Off_Sc%_diff`[i] + nonzero.best.coef[5,2] * games2020predictor$Def_Total_Points_diff[i]
}
```


```{r}
#predict(best.mod,newx=newx)
```



In separate RMD file, also tried refitting model with additional DVOA variables, which seemed to improve performance of predictive models.







